Toward a secure TCP/IP stack --- Technical report

### 2020-07-21 report/report.tex
Typo/English review

    TCP is a widely used network protocol to communicate in the Internet as
    it is used by the HTTP protocol, the FTP protocol, and so many others.
    Ensuring the safety of the TCP/IP stack is essential for the safety of
    a lot of machines. While a lot of work has been done on higher level
    protocols, such as SSL/TLS which is built on top of TCP and designed
    to provide a security layer to TCP with for example the work done in
    miTLS~\cite{bhargavan2013implementing}, or in lower level with the work
    on RecordFlux to safely parse data segment~\cite{Reiher_2020}, nothing has
    been done to make a secure TCP implementation. Nonetheless, TLS security
    can only be ensured if the underlying TCP implementation is free of bugs.

    CycloneTCP is a library developped by Oryx Embedded for embedded platforms.
    A large number of platforms are supported, and the library can be used
    with a dozen of OS. The code of the library is written in C and contains
    plenty protocols from low-level network layers like IPv4/IPv6, to transport layers
    and application layers like DHCP or HTTP. In particular an implementation of
    the TCP protocol is provided.

    SPARK 2014 is a programming language designed as a subset of Ada that helps in
    making high reliability sofwares by providing powerful static analysis methods.
    SPARK is able to detect uninitialized variables with control flow analysis,
    and can ensure the absence of run-time errors,
    but also, based on SMT-solver, functionnal behavior can be specified for every
    functions and mathematically proved.

@@

    This report is a synthesis of the challenge of verifying a such protocol
    and the solution we found during the internship to go toward a secure protocol.

@@

    \subsection{Why verifying a TCP stack?}

    As mentionned in the introduction, TCP protocol is the base of a lot of other protocols,
    in particular SSL/TLS. These higher protocols cannot be secure if the underlying TCP protocol
    has bugs. An incorrect implementation of the TCP protocol can lead to a crash
    in the communication if two machines cannot communicate together, error in
    retransmissions, or even security failures. CycloneTCP is designed for embedded
    code, and a failure in the code can be critical.

@@

    A socket is nothing more than a data structures that contains the connection
    information, like for example the local or the remote IP address as well as
    the state of the connection and other information require by the TCP protocol.
    In Ada, a socket will be represented by a pointer to a record type:

@@

    TCP is defined by the norm RFC 793 \cite{rfc793} in a high level language.
    TCP is a realiable, ordered and error-checked connection oriented protocol.
    It means that two computers have to open a connection first before sending data.
    The connection is closed once all the data has been transmitted.
    The flow of a connection (if no error happens) contains three main steps:
    \begin{enumerate}
        \item Opening the connection,
        \item Sending \& receiving the data,
        \item Closing the connection.
    \end{enumerate}

@@

    As it can be seen in the state machine diagram in figure \ref{fig:TCPAutomaton},
    a socket can take different states that are
    subject to change when a message (or a segment) is sent or received or when an action
    is performed by the user. A label $A/B$ on an edge in the graph refers to a stimuli performed on the socket
    (the reception of a message, a user action...) for $A$ and a flag sent for $B$.
    More precisely, for $A$, a text in italic shape refers to an action performed by
    the user (a call of a user function) or when it's a timeout the action is automatically
    performed by a timer, and a ACK, SYN or FIN refers to a flag contained in a received message.
    The flags in $B$ refers to the flags sent in response of the stimuli $A$.
    Without giving unnecessary details on the format of a TCP header, we can
    describe some flags that can be contained in a TCP header in order to help the understanding
    of the figure \ref{fig:TCPAutomaton}:
    \begin{description}
        \item[ACK] Acknowledgment field significant. The last message received by the sender
        is acknowledge. A field of the header contains the number of this segment.
        \item[SYN] Synchronize sequence number. This flag is sent in order to establish
        a connection.
        \item[FIN] No more data from sender. This flag is sent in order to close the
        connection.
    \end{description}

@@

    For more details on how two TCP can establish or close a connection, the reader can refer
    to the TCP norm where pleinty of scenario are described and explained, and more
    details are given on the state and segments that can be sent.

@@

    \paragraph{User calls}
    User calls refers to functions that can be called by the user to control the connection,
    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between state of
    the connection can happen during the call of these function since there are made to control
    the connection.

    \paragraph{Arriving segments}
    In this threads, the received segments are processed. The corresponding messages are sent back.
    Transitions between states can happen here when a message is received. For example,
    this is the case when a message containing a SYN is received when the socket is in
    the LISTEN state. It will automatically send a SYN + ACK in response, and change
    its state for SYN-RECEIVED.

    \paragraph{Timers}
    The timers control the timeouts like the retransmission timeout to retransmit a message, or
    the time-wait timeout to close the connection after an amont of time. Then, transitions can also
    happen in this threads.

    To sum up, a single socket can be seen as a shared structure over multiples threads, that can all
    manipulate it and change the value of its fields. The schema in figure \ref{fig:TCPConcurrency} illustrates that.

@@

    \begin{figure}
        \begin{tikzpicture}
            \node [draw, ellipse] (S) at (2, 2) {Socket};
            \node [draw, text width=2.5cm, anchor=north] (T1) at (-2,0) {\textbf{TIMER TASK} \\
                                                    Timeout when need};
            \node [draw, text width=2.5cm, anchor=north] (T2) at (2,0) {\textbf{USER TASK} \\
                                                    Controling the connection on the user side};
            \node [draw, text width=2.5cm, anchor=north] (T3) at (6, 0) {\textbf{RECEIVED TASK} \\
                                                    Processing received messages};
            \draw[->] (T1) -- (S);
            \draw[->] (T2) -- (S);
            \draw[->] (T3) -- (S);
            \draw[<->, dashed] (T1) -- (T2);
            \draw[<->, dashed] (T2) -- (T3);

@@

    The TCP norm defines a lot of behavior that could be verified by automatic proof.
    Here is a non exhaustive list of what could be done:
    \begin{itemize}
        \item Verification of the transitions between the states.
        \item Integrity of the messages sent, \textit{i.e.} we could check that the sent
        message contains the correct flag and the correct sequence and acknowledgement numbers.
        \item Integrity of the messages received, \textit{i.e.} we could check that the
        messages received are correctly processed in regard of the flags they contain.
        \item Functionnal correctness of the user functions in regard to their specifications.
    \end{itemize}

    Some choices have had to be done in regard to the short time devolved to the internship, and
    not every aspect of the TCP norm has been proved yet. However, some aspects have been selected
    for the challenge or the importance they represent. A list of these aspects is presented
    here and the choices will be motivated in the next subsections. First of all, we have wanted
    to improve the interface of the functions. The other point was to improved the safety of the
    protocol itself.

    \subsubsection{Improve functions interface}

    One of the main problems pointed by the main writter of the CycloneTCP library is that users
    do not correctly use the API supplied. As a result, some calls of
    functions can be incorrect, either because the arguments are incorrect, or because the
    return code of the previous functions has not been checked, which drive to an
    incorrect code.

    \subsubsection{Respect to the protocol}

    To ensure the safety of the library, it is necessary to ensure that the functions really
    do what they are supposed to do. It is a part of the job that has been done. More
    especially we have wanted to check if the state machine is respected by all the
    user functions in the transition they do. We have also tried to ensure the correctness
    of the function called regarding to the TCP state of the socket before the call
    to respect the specifications given in the norm.

@@

    The multiples tasks and the changes that can happen at every moment in one
    task or another make the verification hard. Multiples interactions exist
    between the multiples threads: synchronous and asynchronous.
    All these interraction have to be considered if we want to write correct contracts
    in regard of what could be done in other threads.

    As SPARK does not have a native mode to deal with concurrency, we have to
    model these different interactions by hands by writing assertions that
    modify the state of the socket as another thread could do it. In particular
    the problem has been encountered when we tried to write contracts for the correctness
    of the user functions. More details and explainations will be given in the following
    section, where we will investigate the solution found to these technical problems.

@@

    This part mainly focussed on the high level user functions, the socket functions.
    The socket functions are the one manipulated by the user to perform operations on
    the network. They are located in the files \texttt{socket\_interface.ad(b|s)}.

    It is clear that there exist an order on which the functions have to be called.
    This order can be determined by the norm, and must respect the order given by the
    graph in figure \ref{fig:TCPAutomaton}.
    Then, we want post- and pre-conditions to model a partial order on the calls of the function.
    If two functions $f_1$ and $f_2$ are ordered such that $f_1 \preceq f_2$ where $\preceq$ is
    a relation over the oder of which functions have to be called, then we want that the
    post-condition of $f_1$ implies the precondition of $f_2$, \textit{i.e.} $\Pre(f_2) \subseteq \Post(f_1)$.

    We will give an example over functions \lstinline[language=Ada]{Socket_Connect} and
    \lstinline[language=Ada]{Socket_Send}. The function \lstinline[language=Ada]{Socket_Connect}
    try to connect to a distant TCP. If the connection succeeds, the remote IP address of the distant
    TCP is set in the field \lstinline[language=Ada]{S_Remote_Ip_Addr} of the \lstinline[language=Ada]{Sock}
    structure. When a user call the function \lstinline[language=Ada]{Socket_Send}, we want to ensure that
    the connection has already been established. This is why a precondition of this function is
    \lstinline[language=Ada]{Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)}.

@@

procedure Socket_Send
    (Sock    : in out Not_Null_Socket;
     Data    : in     Send_Buffer;
     Written :    out Natural;
     Flags   :        Socket_Flags;
     Error   :    out Error_T)
  with
   Pre  =>
     Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)
        \end{lstlisting}
        \caption{An example of how functions calls can be ordered by Pre- and Post-conditions.}
        \label{code:FunOrdered}
    \end{figure}

@@

    An observation made by Clément Zeller, the main programmer of the library is that customers
    do not always think to check the return code of the socket user functions, that can fail.
    If the return code is not checked, some assumption on the result cannot be done.

    The post-conditions as we have written them in SPARK, ensure that the return code is checked
    before continuating processing. Figure \ref{code:FunOrdered} shows this mechanism for the
    procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition filters the cases
    where \lstinline{Error} is \lstinline{NO_ERROR} and where \lstinline{Error} takes another value.
    As a result, a user who would write an incorrect code such as
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    would be warned by gnatprove with a message such as
    \begin{verbatim}
    medium: precondition might fail.
    \end{verbatim}
    whereas the following code is correct
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    if Error /= NO_ERROR then
        return;
    end if;
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}

@@

    The aim of this part is to explain how we have done to check that the transitions done
    in a function respect the TCP automaton. In particular, we were interested in verifying
    the high level user functions. These functions are not the one in which most of the
    transitions are done. Even, the verification of these functions is important in order
    to guarantee the safety of the whole library. The functions of interest are located in
    the file \texttt{tcp\_interface.ad(b|s)}.

    In order to verify the transitions we have read the norm to extract informations about
    the allowed transitions, and we have added contracts in the function
    \lstinline[language=Ada]{tcpChangeState}. Since all the transitions are made through
    this function, an incorrect transition will be warned by SPARK.

    As mentionned, since most of transitions are done outside the user functions, a big part
    of this work was to consider other transitions in other threads and we present how we have
    done it in the following sections.

@@

    To ensure the safety of our library, we need to consider all what can happen everywhere in
    the library. All the functions are protected by mutex, which means that only one operation
    can be performed at the same time on a socket. Interractions must be considered at two
    locations: between the function calls, when the mutex is released and during the function
    call, when the program waits for an event. We will explore in the sections \ref{section:asynchronous}
    and \ref{section:synchronous} how we have dealt with this two different concurrent mechanisms.

    Before everything else, let's see where the interferences mainly come from, and how we have
    done to ensure that all have been considered.

    \subsubsection{Process segments functions}

@@

    The file \texttt{tcp\_fsm.c} is in charge of processing the incoming segments. The reader is
    strongly encouraged to have a look at this file. The main function of this file
    \texttt{tcpProcessSegment} looks for the socket corresponding to the received segment, and then
    according to the TCP state of the socket, process the segment as expected by calling one of
    the function \texttt{tcpState<StateName>}. The functions familly \texttt{tcpState<StateName>}
    check the information contained in the segment and can perform a change of state in the socket
    struct depending on the flags received.

    In these functions, a change of state can be performed, as mentioned, thanks to the function
    \texttt{tcpChangeState}. To summarize what is done when a message is received, rather than
    reading the code and locate all the calls to functions \texttt{tcpChangeState}, we have used
    Klee\footnote{\url{https://klee.github.io}}. It helped to find behaviors that had not been
    imagined at first. Anyway, we cannot state that all the work done with Klee is complete, and
    we would need to rewrite this part in SPARK to have a code formally proved.

    All what is related to Klee is present in the folder \texttt{klee/} and can be compiled and run
    thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
    incomming message, put the socket in the desired state and call the desired function to see what
    final states can be obtain.

    Finally, all the results found by Klee has been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
    This function is essential for all the verification of concurrent parts and is reused everywhere
    the concurrency is involved to compute possible interferences. It follows that the safety of the code rely on the
    confidence we have in this function.

@@

    Between the user functions calls, segments can be received and these receptions can drive to
    changes of the state of the socket. Between two function calls, an infinite number of segments
    can be potentially receive. The reception of one or zero segment is modeled by the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} in term of change of states. Then we have
    to consider the iteration of \lstinline[language=Ada]{Tcp_Process_One_Segment} in order to
    compute the result of the reception of multiple message.

    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
    that can be performed by the reception of a message and its automatic response mechanism.
    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without user action. Since
    we only consider the transitive closure in term of states, we can significantly reduce the number
    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
    small loops to obtain more precise results.

@@

    In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all the
    possible that all the possible input states are have been considered.

    \subsubsection{Concurrency: Synchronous exchange}
    \label{section:synchronous}

    The other mechanism to deal with concurrency is probably more difficult to apprehend and the
    use of the function \lstinline[language=Ada]{Tcp_Process_One_Segment} is even more noticeable.

    The C code contains a function \lstinline[language=C]{tcpWaitForEvents} in the file \texttt{tcp\_misc.c}
    that checks if the event is true when the function is called, by calling the function
    \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
    called, then the mutex that was previously locked is released until the waited event happens.
    Then, everything can happen meanwhile.

@@

    When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} that reuse
    \lstinline[language=Ada]{Tcp_Process_One_Segment} to compute the most precise set of possible
    final states after the wait. \textit{(This is done in the function \lstinline[language=Ada]{Tcp_Wait_For_Events_Proof},
    the function dedicated to proof in file \texttt{tcp\_misc\_binding.adb}.)}

@@

    We can compute precisely the states reach for each event waited thanks to the fact that SPARK
    unrolls loops.

>>>>

    TCP is a widely used network protocol to communicate in the Internet as
    it is used by the HTTP protocol, the FTP protocol, and so many others.
    Ensuring the safety of the TCP/IP stack is essential for the safety of
    a lot of machines. While a lot of work has been done on formally verifying higher level
    protocols, such as SSL/TLS which is built on top of TCP and designed
    to provide a security layer to TCP with for example the work done in
    miTLS~\cite{bhargavan2013implementing}, or at a lower level with the work
    on RecordFlux to safely parse data segments~\cite{Reiher_2020}, nothing has
    been done to formally verify a secure TCP implementation. However, TLS security
    can only be ensured if the underlying TCP implementation is free of bugs.

    CycloneTCP is a library developped by the company Oryx Embedded for embedded platforms.
    A large number of platforms are supported, and the library can be used
    with a dozen of OS. The code of the library is written in C and implements
    many protocols from low-level network layers like IPv4/IPv6, to transport layers
    and application layers like DHCP or HTTP. In particular an implementation of
    the TCP protocol is provided.

    SPARK 2014 is a programming language designed as a subset of Ada that helps in
    making high reliability sofware by providing powerful static analysis methods.
    SPARK is able to detect uninitialized variables with control flow analysis,
    and can ensure the absence of run-time errors,
    but also, based on SMT-solver, functional behavior can be specified for every
    function and mathematically proved.

@@

    This report is a synthesis of the challenge of verifying such a protocol
    and the solution we found during the internship to go towards a secure protocol.

@@

    \subsection{Why verify a TCP stack?}

    As mentioned in the introduction, TCP protocol is the base of a lot of other protocols,
    in particular SSL/TLS. These higher protocols cannot be secure if the underlying TCP protocol
    has bugs. An incorrect implementation of the TCP protocol can lead to a crash
    in the communication if two machines cannot communicate together, to an error in
    retransmissions, or even to security failures. CycloneTCP is designed for embedded
    code, and a failure in the code can be critical.

@@

    A socket is nothing more than a data structure that contains the connection
    information, like for example the local or the remote IP address as well as
    the state of the connection and other information required by the TCP protocol.
    In Ada, a socket will be represented by a pointer to a record type:

@@

    TCP is defined by the norm RFC 793 \cite{rfc793} in a high level language.
    TCP is a reliable, ordered and error-checked connection oriented protocol.
    It means that two computers have to open a connection first before sending data.
    The connection is closed once all the data has been transmitted.
    The flow of a connection (if no error happens) contains three main steps:
    \begin{enumerate}
        \item Opening the connection,
        \item Sending \& receiving the data,
        \item Closing the connection.
    \end{enumerate}

@@

    As it can be seen in the state machine diagram in figure \ref{fig:TCPAutomaton},
    a socket can take different states that are
    subject to change when a message (or a segment) is sent or received or when an action
    is performed by the user. A label $A/B$ on an edge in the graph refers to a stimulus performed on the socket
    (the reception of a message, a user action...) for $A$ and a flag sent for $B$.
    More precisely, for $A$, a text in italic shape refers to an action performed by
    the user (a call of a user function) or when it's a timeout the action is automatically
    performed by a timer, and a ACK, SYN or FIN refers to a flag contained in a received message.
    The flags in $B$ refer to the flags sent in response of the stimulus by $A$.
    Without giving unnecessary details on the format of a TCP header, we can
    describe some flags that can be contained in a TCP header in order to help the understanding
    of figure \ref{fig:TCPAutomaton}:
    \begin{description}
        \item[ACK] Acknowledgment field significant. The last message received by the sender
        is acknowledged. A field of the header contains the number of this segment.
        \item[SYN] Synchronize sequence number. This flag is sent in order to establish
        a connection.
        \item[FIN] No more data from sender. This flag is sent in order to close the
        connection.
    \end{description}

@@

    For more details on how two TCP can establish or close a connection, the reader can refer
    to the TCP norm where many scenarios are described and explained, and more
    details are given on the state and segments that can be sent.

@@

    \paragraph{User calls}
    User calls refer to functions that can be called by the user to control the connection,
    send, or received data: OPEN, CLOSE, ABORT, SEND, RECEIVED. Transition between states of
    the connection can happen during the call of these functions since they are intended to control
    the connection.

    \paragraph{Arriving segments}
    In this thread, the received segments are processed. The corresponding messages are sent back.
    Transitions between states can happen here when a message is received. For example,
    this is the case when a message containing a SYN is received when the socket is in
    the LISTEN state. It will automatically send a SYN + ACK in response, and change
    its state for SYN-RECEIVED.

    \paragraph{Timers}
    The timers control the timeouts like the retransmission timeout to retransmit a message, or
    the time-wait timeout to close the connection after an amount of time. Then, transitions can also
    happen in this thread.

    To sum up, a single socket can be seen as a shared structure over multiple threads, that can all
    manipulate it and change the value of its fields. The schema in figure \ref{fig:TCPConcurrency} illustrates that.

@@

    \begin{figure}
        \begin{tikzpicture}
            \node [draw, ellipse] (S) at (2, 2) {Socket};
            \node [draw, text width=2.5cm, anchor=north] (T1) at (-2,0) {\textbf{TIMER TASK} \\
                                                    Timeout when need};
            \node [draw, text width=2.5cm, anchor=north] (T2) at (2,0) {\textbf{USER TASK} \\
                                                    Controlling the connection on the user side};
            \node [draw, text width=2.5cm, anchor=north] (T3) at (6, 0) {\textbf{RECEIVED TASK} \\
                                                    Processing received messages};
            \draw[->] (T1) -- (S);
            \draw[->] (T2) -- (S);
            \draw[->] (T3) -- (S);
            \draw[<->, dashed] (T1) -- (T2);
            \draw[<->, dashed] (T2) -- (T3);

@@

    The TCP norm defines a lot of behavior that could be verified by automatic proof.
    Here is a non exhaustive list of what could be done:
    \begin{itemize}
        \item Verification of the transitions between the states.
        \item Integrity of the messages sent, \textit{i.e.} we could check that the sent
        message contains the correct flag and the correct sequence and acknowledgement numbers.
        \item Integrity of the messages received, \textit{i.e.} we could check that the
        messages received are correctly processed in regard of the flags they contain.
        \item Functional correctness of the user functions in regard to their specifications.
    \end{itemize}

    Some choices have had to be done in regard to the short time devolved to the internship, and
    not every aspect of the TCP norm has been proved yet. However, some aspects have been selected
    for the challenge or the importance they represent. A list of these aspects is presented
    here and the choices will be motivated in the next subsections. First of all, we have wanted
    to improve the interface of the functions. The other point was to improve the safety of the
    protocol itself.

    \subsubsection{Improve the functional interface}

    One of the main problems pointed to by the main author of the CycloneTCP library is that users
    do not correctly use the API supplied. As a result, some function
    calls can be incorrect, either because the arguments are incorrect, or because the
    return code of the previous calls has not been checked, which lead to an
    incorrect code.

    \subsubsection{Respect of the protocol}

    To ensure the safety of the library, it is necessary to ensure that the functions really
    do what they are supposed to do. It is a part of the job that has been done. More
    especially we have wanted to check if the state machine is respected by all the
    user functions in the transition they do. We have also tried to ensure the correctness
    of the functions called regarding the TCP state of the socket before the call
    to respect the specifications given in the norm.

@@

    The multiples tasks and the changes that can happen at every moment in one
    task or another make the verification hard. Multiples interactions exist
    between the multiples threads: synchronous and asynchronous.
    All these interactions have to be considered if we want to write correct contracts
    in regard of what could be done in other threads.

    As SPARK does not have a native mode to deal precisely with interactions related to concurrency, we have to
    model these different interactions by hands by writing assertions that
    modify the state of the socket as another thread could do it. In particular
    the problem has been encountered when we tried to write contracts for the correctness
    of the user functions. More details and explanations will be given in the following
    section, where we will investigate the solution found to these technical problems.

@@

    This part mainly focusses on the high level user functions, the socket functions.
    The socket functions are the one called by the user to perform operations on
    the network. They are located in the files \texttt{socket\_interface.ad(b|s)}.

    It is clear that there exists an order in which the functions have to be called.
    This order can be determined by the norm, and must respect the order given by the
    graph in figure \ref{fig:TCPAutomaton}.
    Then, we want post- and pre-conditions to model a partial order on the calls of the function.
    If two functions $f_1$ and $f_2$ are ordered such that $f_1 \preceq f_2$ where $\preceq$ is
    a relation over the order in which functions have to be called, then we want that the
    post-condition of $f_1$ implies the precondition of $f_2$, \textit{i.e.} $\Pre(f_2) \subseteq \Post(f_1)$.

    We will give an example over functions \lstinline[language=Ada]{Socket_Connect} and
    \lstinline[language=Ada]{Socket_Send}. The function \lstinline[language=Ada]{Socket_Connect}
    tries to connect to a distant TCP. If the connection succeeds, the remote IP address of the distant
    TCP is set in the field \lstinline[language=Ada]{S_Remote_Ip_Addr} of the \lstinline[language=Ada]{Sock}
    structure. When a user calls the function \lstinline[language=Ada]{Socket_Send}, we want to ensure that
    the connection has already been established. This is why a precondition of this function is
    \lstinline[language=Ada]{Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)}. This precondition will only
    be proved if \lstinline[language=Ada]{Socket_Connect} has been previously called.

@@

procedure Socket_Send
    (Sock    : in out Not_Null_Socket;
     Data    : in     Send_Buffer;
     Written :    out Natural;
     Flags   :        Socket_Flags;
     Error   :    out Error_T)
  with
   Pre  =>
     Is_Initialized_Ip(Sock.S_Remote_Ip_Addr)
        \end{lstlisting}
        \caption{An example of how function calls can be ordered by Pre- and Post-conditions.}
        \label{code:FunOrdered}
    \end{figure}

@@

    An observation made by Clément Zeller, the main programmer of the library is that customers
    do not always think to check the return code of the socket user functions, to deal with the
    possibility that the call fails.
    If the return code is not checked, some assumption on the result cannot be done.

    The post-conditions as we have written them in SPARK, ensure that the return code is checked
    before processing continues. Figure \ref{code:FunOrdered} shows this mechanism for the
    procedure \lstinline[language=Ada]{Socket_Connect}. The post-condition distinguishes the cases
    where \lstinline{Error} is \lstinline{NO_ERROR} and where \lstinline{Error} takes another value.
    As a result, a user who would write an incorrect code such as
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    would be warned by gnatprove with a message such as
    \begin{verbatim}
    medium: precondition might fail.
    \end{verbatim}
    whereas the following code is correct
    \begin{lstlisting}[language=Ada]
    Socket_Connect (Sock, Remote_Ip_Addr, Port, Error);
    if Error /= NO_ERROR then
        return;
    end if;
    Socket_Send (Sock, Data, Written, Flags, Error);
    \end{lstlisting}

    So in the end, the precondition of \lstinline[language=Ada]{Socket_Send} will only
    be proved if \lstinline[language=Ada]{Socket_Connect} has been previously called, and the code
    checks after that call that \lstinline[language=Ada]{Error = NO_ERROR}, and the socket
    \lstinline[language=Ada]{Sock} has not been modified.

@@

    The aim of this part is to explain how we check that the transitions done
    in a function respect the TCP automaton. In particular, we are interested in verifying
    the high level user functions. These functions are not the one in which most of the
    transitions are done. Still, the verification of these functions is important in order
    to guarantee the safety of the whole library. The functions of interest are located in
    the file \texttt{tcp\_interface.ad(b|s)}.

    In order to verify the transitions we have read the norm to extract informations about
    the allowed transitions, and we have added contracts in the function
    \lstinline[language=Ada]{tcpChangeState}. Since all the transitions are made through
    this function, an incorrect transition will lead to a message by SPARK.

    As mentioned, since most transitions are done outside the user functions, a big part
    of this work was to consider other transitions in other threads and we present how we have
    done it in the following sections.

@@

    To ensure the safety of our library, we need to consider all what can happen everywhere in
    the library. All the functions are protected by mutex, which means that only one operation
    can be performed at the same time on a socket. Interactions must be considered at two
    locations: between the function calls, when the mutex is released and during the function
    call, when the program waits for an event. We will explore in the sections \ref{section:asynchronous}
    and \ref{section:synchronous} how we have dealt with these two different concurrent mechanisms.

    Before everything else, let's see where the interferences mainly come from, and how we have
    ensured that all have been considered.

    \subsubsection{Functions that process segments}

@@

    The file \texttt{tcp\_fsm.c} is in charge of processing the incoming segments. The reader is
    strongly encouraged to have a look at this file. The main function of this file
    \texttt{tcpProcessSegment} looks for the socket corresponding to the received segment, and then
    according to the TCP state of the socket, processes the segment as expected by calling one of
    the function \texttt{tcpState<StateName>}. The family of functions \texttt{tcpState<StateName>}
    check the information contained in the segment and can perform a change of state in the socket
    struct depending on the flags received.

    In these functions, a change of state can be performed, as mentioned, thanks to the function
    \texttt{tcpChangeState}. To summarize what is done when a message is received, rather than
    reading the code and locate all the calls to functions \texttt{tcpChangeState}, we have used
    KLEE\footnote{\url{https://klee.github.io}}. It helped to find behaviors that we had not 
    imagined at first. Anyway, we cannot state that all the work done with KLEE is complete, and
    we would need to rewrite this part in SPARK to have a code formally proved.

    All the work related to KLEE is present in the folder \texttt{klee/} and can be compiled and run
    thanks to the makefile to reproduce the results. Roughtly, what is done is creating a random
    incoming message, put the socket in the desired state and call the desired function to see what
    final states can be obtain.

    Finally, all the results found by KLEE have been reported as a post-condition of the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
    This function is essential for all the verification of concurrent parts and is reused everywhere
    the concurrency is involved to compute possible interferences. It follows that the safety of the code rely on the
    confidence we have in this function.

@@

    Between the user function calls, segments can be received and these receptions can lead to
    changes of the state of the socket. Between two function calls, an infinite number of segments
    can be potentially received. The reception of one or zero segments is modeled by the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} in terms of change of states. Then we have
    to consider the iteration of \lstinline[language=Ada]{Tcp_Process_One_Segment} in order to
    compute the result of the reception of multiple messages.

    Let $\rightarrow$ be the transition function in the TCP automaton restricted to the transitions
    that can be performed by the reception of a message and its automatic response mechanism.
    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without user action. Since
    we only consider the transitive closure in term of states, we can significantly reduce the number
    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
    It follows that algorithm \ref{algo:tcpProcessSegment} is good enough to compute the transitive closure of the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
    small loops to obtain more precise results.

@@

    In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all 
    possible input states have been considered.

    \subsubsection{Concurrency: synchronous exchange}
    \label{section:synchronous}

    The other mechanism to deal with concurrency is probably more difficult to comprehend and the
    use of the function \lstinline[language=Ada]{Tcp_Process_One_Segment} is even more noticeable.

    The C code contains a function \lstinline[language=C]{tcpWaitForEvents} in the file \texttt{tcp\_misc.c}
    that checks if the event is true when the function is called, by calling the function
    \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
    called, then the mutex that was previously locked is released until the expected event happens.
    Then, everything can happen meanwhile.

@@

    When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} that reuses
    \lstinline[language=Ada]{Tcp_Process_One_Segment} to compute the most precise set of possible
    final states after the wait. \textit{(This is done in the function \lstinline[language=Ada]{Tcp_Wait_For_Events_Proof},
    the function dedicated to proof in file \texttt{tcp\_misc\_binding.adb}.)}

@@

    We can compute precisely the states reached for each expected event thanks to the fact that SPARK
    unrolls loops.
